---
title: "lab5_codealong"
author: "Genevieve Chiong"
date: "2/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(broom)

# Time series packages
library(tsibble)
library(feasts)
library(fable)

# Spatial data packages
library(sf)
library(tmap)
```

# Part 1: Time series exploration and forecasting 
```{r}
energy <- read_csv(here("data", "energy.csv"))
```

```{r}
energy_ts <- energy %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date)
```

## Raw data graph
```{r}
ggplot(data = energy_ts, aes(x = date, y = res_total)) +
  geom_line() + 
  labs(y = "Residential energy consumption \n (Trillion BTU)")
```

## Seasonplot 
```{r}
energy_ts %>% 
  gg_season(y = res_total) +
  theme_minimal() +
  labs(x = "month",
       y = "residential energy consumption (trillion BTU)")
```

- The highest residential energy usage is around December / January / February
- There is a secondary peak around July & August (that’s the repeated secondary peak we see in the original time series graph)
- We can also see that the prevalence of that second peak has been increasing over the course of the time series: in 1973 (orange) there was hardly any summer peak. In more recent years (blue/magenta) that peak is much more prominent.

## Subseries plot 

```{r}
energy_ts %>% gg_subseries(res_total)
```

- Our takeaway here is similar: there is clear seasonality (higher values in winter months), with an increasingly evident second peak in June/July/August. This reinforces our takeaways from the raw data and seasonplots.

## Decomposition 

```{r}
# Find STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))

# View the components
components (dcmp)

# Visualize the decomposed components
components(dcmp) %>% autoplot() + 
  theme_minimal() 

```

## Autocorrelation function (ACF) 

```{r}
energy_ts %>% 
  ACF(res_total) %>% 
  autoplot() 
```

- We see that observations separated by 12 months are the most highly correlated, reflecting strong seasonality we see in all of our other exploratory visualizations.

## Holt-Winters exponential smoothing 

```{r}
# Create the model: 
energy_fit <- energy_ts %>% 
  model(
    ets = ETS(res_total ~ season("M"))
  )

# Forecast using the model 10 years into the future: 
energy_forecast <- energy_fit %>% 
  forecast(h = "10 years")

# Plot just the forecasted values (with 80% and 95% CIs): 
energy_forecast %>% 
  autoplot() 
```

```{r}
# Or plot it added to the original data: 

energy_forecast %>% 
  autoplot(energy_ts)
```

## Assessing residuals 

```{r}
# Append the predicted values (and residuals) to original energy data 
energy_predicted <- broom::augment(energy_fit)

# Use View(energy_predicted) to see the resulting data frame
```

```{r}
# Plot the actual energy values (res_total), and the predicted values (stored as .fitted) atop them
ggplot(data = energy_predicted) + 
  geom_line(aes(x = date, y = res_total)) + 
  geom_line(aes(x = date, y = .fitted), color = "red")
```

Residuals should be uncorrelated, centered at 0, and ideally normally distributed. 

```{r}
ggplot(data = energy_predicted, 
       aes(x = .resid)) + 
  geom_histogram()
```


# Part 2: Spatial data wrangling, visualization, and a variogram 

A. California county outlines (polygons) 

```{r}
# Read in the CA county shapefile 

ca_counties <- read_sf(here("data","ca_counties","CA_Counties_TIGER2016.shp"))

ca_subset <- ca_counties %>% 
  select(NAME, ALAND) %>% 
  rename(county_name = NAME, land_area = ALAND)

# Check and set the CRS 

ca_subset %>% st_crs() 
```

```{r}
# Plot the California counties using geom_sf()

ggplot(data = ca_subset) + 
  geom_sf(aes(fill = land_area),
          color = "white", 
          size = 0.1) +
  theme_void() + 
  scale_fill_gradientn(colors = c("cyan", "blue", "purple"))
```

B. Invasive red sesbania records (spatial points) 

```{r}
# Read in the data 
sesbania <- read_sf(here("data", "red_sesbania", "ds80.shp"))

# Check the CRS: 
sesbania %>% st_crs()
```

```{r}
# Update the CRS so it matches the CA counties CRS using st_transform()
sesbania <- st_transform(sesbania, 3857)

# Then check it: 
sesbania %>% st_crs()
```

```{r}
# Plot them together! 
  
ggplot() + 
  geom_sf(data = ca_subset) +
  geom_sf(data = sesbania, size = 1, color = "red")
```

```{r}
# Find the count of red sesbania observed locations in this dataset by county 

# Spatial joins 
ca_sesbania <- ca_subset %>% 
  st_join(sesbania) 

sesbania_counts <- ca_sesbania %>% 
  count(county_name)
```

```{r}
# Chloropleth using the number of records for red sesbania as the fill color 

ggplot(data = sesbania_counts) + 
  geom_sf(aes(fill = n), color = "white", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgray", "orange", "red")) +
  theme_minimal() +
  labs(fill = "Number of S. punicea records")
```

```{r}
# Suset of sesbania point locations only in Solano County
solano_sesbania <- sesbania %>% 
  filter(COUNTY == "Solano") 

# Onlykeep Solano polygon from California County data 
solano <- ca_subset %>% 
  filter(county_name == "Solano")

ggplot()+
  geom_sf(data = solano) +
  geom_sf(data = solano_sesbania)
```

C. Making an interactive map with {tmap}

```{r}
# Set the viewing mode to "interactive": 
tmap_mode(mode = "view")

# Then make a map (with the polygon fill color updated by variable 'land_area', updating the color palette to "BuGn"), then add another shape layer for the sesbania records (added as dots): 

tm_shape(ca_subset) + 
  tm_fill("land_area", palette = "BuGn") +
  tm_shape(sesbania) + 
  tm_dots()
```




